---
title: "Introduction to fia"
author: "Jose Eduardo Meireles"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to fia}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
  
 ## Installation
 
The easiest way to get the `fia` package is to install from the bitbuket repository.
  
```{r eval=FALSE}
library("devtools")
devtools::install_bitbucket("meireles/fia", quiet = TRUE)
```

## Usage

For now, the main goals of the `fia` package is to programatically download FIA data, which lives lives in the FIADB webpage. 

First load the `fia`  package and then checl out the default FIA url.

```{r}

# Load the `fia` library
library("fia")

# And check out the packages' default FIA url using
fia::url_fia()
```


The FIA dataset is complex and composed of many different tables. There are two main types of tables:

- Reference: These tables are prefixed `REF_`. They hold. as the name says, reference data such as species names (`REF_SPECIES`) and forest types (`REF_FOREST_TYPE`).

- Data: These are tables with plot data (collected at the sites). For example, the individual tree data in `TREE` or geometry of each plot in `PLOTGEOM`. Individual tables can be obtained at the state level or for all states toguether.

To find out what tables are available in the FIA, use `list_available_tables`. 

```{r}
# Rentrieve table names from the web
tab = fia::list_available_tables()

# Kinds of tables available...
print(tab)

# So you can ckeck out the names of reference tables
head(tab$reference_table_names)

# Or plot level data tables and state names
head(tab$data_table_names)
head(tab$data_states)

# As well as many the raw-ish matrices with table file names, modicication dates, etc...
head(tab$reference_tables)
```

Once you figured what data you want to grab, all you have to do is download it:

```{r eval=TRUE}
# Pick tables to download. Chosen in this case from tab$reference_table_names and tab$data_table_names
my_table_names =  c("REF_SPECIES",  # a reference table
                    "PLOT")         # a plot data table

# Then choose the states you're interest in. See the pssibilities in tab$data_states.
my_states     = c("NC", "MN")

# Alternativelly, choose all states in the dataset. Will retrieve data in tab$data_tables_states_combined
my_states_2   = "ALL"

# You also must provide a directory where all the files will be dumped in.
my_output_dir = "data_test/fia/"

# Start download, will probably take a while...
fia::download_fia_tables(table_names           = my_table_names,
                         states                = my_states,
                         destination_dir       = my_output_dir,
                         list_available_tables = tab)           # the object returned by fia::list_available_tables()
```
