---
title: "Introduction to fia"
author: "Jose Eduardo Meireles"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to fia}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
  
  The easiest way to get the `fia` package is to install from the bitbuket repository.
  
```{r eval=FALSE}
library("devtools")
devtools::install_bitbucket("meireles/fia", quiet = TRUE)
```

Now you can just load the package.

```{r}
library("fia")
```

All of the data lives in the FIADB webpage. You can see the default FIA url, simply run `url_fia()` and check the results.

```{r}
fia::url_fia()
```

The FIA dataset is complex and composed of many different tables. There are two main types of tables, holding:

- Reference Data: For instance, species names in the `REF_SPECIES` table.
- Plot data: For example, the individual tree data in `TREE` or geometry of each plot in `PLOTGEOM`.

In general, to find out what tables are available in the fia using:

```{r}
tab = fia::list_available_tables() # returns a list with the tables and states available in the dataset.

## Just for kicks, check our the header of each object in the list
lapply(tab, head)
```

Once you figured what data you want to grab, go ahead and download it with:

```{r eval=TRUE}

# Pick tables to download. Chosen from tab$reference_table_names and tab$data_table_names
my_table_names =  c("REF_SPECIES",  # a reference table
                    "PLOT")         # a plot data table

# Then choose the states you're interest in. See the pssibilities in tab$data_states.
my_states     = c("NC", "MN")

# Alternativelly, choose all states in the dataset. Will retrieve data in tab$data_tables_states_combined
my_states_2   = "ALL"

# You also must provide a directory where all the files will be dumped in.
my_output_dir = "data_test/fia/"

fia::download_fia_tables(table_names           = my_table_names,
                         states                = my_states,
                         destination_dir       = my_output_dir,
                         list_available_tables = tab)           # the object returned by fia::list_available_tables()
```
