---
title: "Introduction to fia"
author: "Jose Eduardo Meireles"
#date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    keep_md: TRUE
vignette: >
  %\VignetteIndexEntry{Introduction to fia}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

### Installation
 
The easiest way to get the `fia` package is to install from the github repository.

```{r eval=FALSE}
library("devtools")
devtools::install_github("meireles/fia")
```

### Usage

The main goal of the `fia` package is to programatically download FIA data, which lives lives in the FIADB webpage. Use `url_fia()` to see the default FIA url.

```{r}
# Load the `fia` library
library("fia")

# Check out the packages' default FIA url
fia::url_fia()
```

### Finding what data is available

The FIA dataset is complex and composed of many different tables. To find out which tables are available in the FIA, use `list_available_tables`.

```{r}
# Retrieve table names from the web
tab = fia::list_available_tables()

# Types of tables available.
print(tab)
```

There are two main types of tables:

+ **Reference:** Prefixed `REF`. They hold, as the name says, reference data such as species names (`REF_SPECIES`) and forest types (`REF_FOREST_TYPE`). You 'll find them in `reference_table_names`.

+ **Data:** These are tables with plot data (collected at the sites). For example, the individual tree data in `TREE` or geometry of each plot in `PLOTGEOM`. You'll find their names inside `data_table_names`. Note that these individual tables can be obtained:

    + At the **state level:** The state abbreviations are found inside `data_states`. A detailed list of state level datasets is given in `data_tables_by_state`
    
    + For **all states combined:** The detailed list of data combined for all states is given inside `data_tables_states_combined`. 


```{r}
# Check out the names of reference tables
head(tab$reference_table_names)

# or names of states or plot level data tables
head(tab$data_states)
head(tab$data_table_names)

# You can also get a detailed desrciption of the data available in the FIA, e.g.
head(tab$reference_tables)
```


### Downloading FIA tables

Once you figured out what data you want to grab, all you have to do is download it:

```{r eval=TRUE}

# Pick tables to download
# Chosen from tab$reference_table_names and tab$data_table_names
my_table_names =  c("REF_SPECIES",  # a reference table
                    "PLOT")         # a plot data table

# Pick states of interest
# See the pssibilities in tab$data_states
my_states     = c("NC", "MN")

# Alternativelly, choose all states in the dataset
# Thus will retrieve data in tab$data_tables_states_combined
my_states_2   = "ALL"

# You also must provide the directory where the files will be dumped in.
my_output_dir = "data_test/fia"

# Start download!
# Will probably take a while...
fia::download_fia_tables(table_names           = my_table_names,
                         states                = my_states,
                         destination_dir       = my_output_dir,
                         list_available_tables = tab, # returned by `list_available_tables()`
                         table_format          = "zip")
```
